A sheaf--function dictionary for algebraic tori over local fields

### 2013-08-17 shanghai.tex
Working on greenberg of neron section.

  To show that $L_G(f,\psi) \iso (\qcs{L}_0,\mu_0)$ we simply apply the functions--sheaves dictionary
  as follows ({\it cf} \cite{laumon:87a}*{1.1.3} or, ultimately, \cite{deligne:SGA4.5}*{Sommes trigonométriques}).
  First, recall that $L_G(f,\psi) = (L(f,\psi), \mu(f,\psi))$. 
  By  \cite{laumon:87a}*{1.1.3.3}, 
  \[
  \trFrob{(G_{k'}\hookrightarrow G)^*L(f,\psi)} = \trFrob{\qcs{L}} \circ N_{k'/k},
  \]
   for each finite extension $k'/k$.
  On the other hand, 
  \[
  \trFrob{\qcs{L}} \circ N_{k'/k} = \trFrob{(G_{k'}\hookrightarrow G)^*\qcs{L}},
  \]
   by Proposition~\ref{prop:basechange}.
  %Using notation from \cite{laumon:87a}*{1.1}, this means $t_{\bullet L(f,\psi)} = t_{\bullet \qcs{L}_0}$.
  Using \cite{laumon:87a}*{Th\'eor\`eme~1.1.2}, it follows that  $L(f,\psi) \iso \qcs{L}_0$ in $\Loc(G)$.
  It is now clear that $(L(f,\psi),\mu(f,\psi)) \iso (\qcs{L}_0,\mu_0)$ in $\QCb(G)$.
  This completes the proof that $L_G : \QCf(G) \to \QC(G)$ is essentially surjective.
  Since we have already seen (in Section~\ref{sec:finite}) that $L_G$ is full and faithful, 
  it follows that $L_G$ is an equivalence.

@@

Recall that $\Fq$ is a finite field.
Let $K$ be a non-archimedean local field with residue field $\Fq$.
We place no restriction on the characteristic of $K$. 
We write $\OK$ for the ring of integers of $K$ and $\pK$ for the maximal ideal in $\OK$. 
For any positive integer $d$, set $A_d\ceq \OK/\pK^d$.
Let $T$ be an algebraic torus over $K$.
We write $\Hom_{< d}(T(K),\EEx)$ for the group of characters of $T(K)$ depth less than $d$ 
and $\Hom_{<\infty}(T(K),\EEx)$ for the group of characters of $T(K)$ of arbitrary finite depth; one might call these quasicharacters of $T(K)$, expanding on \cite{tate:67a}*{2.3}.

@@

\begin{definition}\label{def:GN}
For each positive integer $d$, set $\TT_{d} \ceq \TT \times_\Spec{\OK} \Spec{A_d}$. Observe that $A_d$ is an Artin ring and $\TT_{d}$ is smooth and locally of finite type over $A_d$.
%
Let $\GN{T}_{d} = \Gr(\TT_d)$ be the group scheme over $\Fq$ produced by applying
the Greenberg functor to $\TT_{d}$; this is the \emph{Greenberg transform of $\TT_d$}.
\end{definition}

See \cite{greenberg:63a} and \cite{bosch-lutkebohmert-reynaud:NeronModels}*{Ch. 9, \S 6} and the first few sections of \cite{stasinski:12a} for background on the Greenberg functor. The fundamental properties of $\GN{T}_d$ that we need are:
 \begin{enumerate}
   \labitem{(GN.0)}{dFq} $\GN{T}_d(\Fq) = \TT(A_d)$, canonically.
  \labitem{(GN.1)}{GNd} $\GN{T}_d$ is a smooth, commutative group scheme, locally of finite type over $\Fq$, and the component group of $\GN{T}_d$ is geometrically finitely generated.
  \end{enumerate}
In fact, $\pi_0(\GN{T}_d) = \pi_0(\TT) \times_{\Spec{\OK}} \Spec{\Fq}$, and the latter is finitely generated by \cite{xarles:93a}.

@@

\begin{proof}
  By \ref{GNd}, we may apply Theorem~\ref{thm:snake} to
  $\GN{T}_d$ which provides a canonical isomorphism
  $\QCiso{\GN{T}_d} \iso \GN{T}_d(\Fq)^*$.
  By \ref{dFq}, $\GN{T}_d(\Fq) = \TT(A_d)$.
  Then we recall that the filtration of $T(K)$ used to
  define depth has the property that $\TT(A_d) \cong T(K) / T(K)_d$
  \cite{yu:03a}*{Prop 5.2}, so characters of $\GN{T}_d(\Fq)$ 
  are in bijection with characters of $T(K)$ vanishing on $T(K)_d$. 
  This shows that $\trFrob{T,d}$ is an isomorphism.

@@

\Clifton{I'm working here, now.}
If $d\leq d'$, then $\pK^{d'} \hookrightarrow \pK^d$ so there is a canonical surjective ring homomorphism $A_{d'} \to A_{d}$. Set $S = \Spec{\OK}$ and $S_d = \Spec{A_d}$. 
\[
\begin{tikzcd}
S_1 \arrow[hook]{r} 
& S_2 \arrow[hook]{r} \cdots 
& S_d \arrow[hook]{r}  
& S_{d'} \arrow[hook]{r} \cdots 
& S
\end{tikzcd}
\]

 then $\GN{T}_d$ is a closed subgroup of $\GN{T}_{d'}$,
so these smooth commutative group schemes form a directed system.
\[
\begin{tikzcd}
\GN{T}_1 
& \arrow[two heads]{l} \GN{T}_2 
& \arrow[two heads]{l} \cdots 
& \arrow[two heads]{l} \GN{T}_d 
& \arrow[two heads]{l} \GN{T}_{d'} 
& \arrow[two heads]{l} \cdots
\end{tikzcd}
\]

>>>>

  To show that $L_G(f,\psi) \iso (\qcs{L}_0,\mu_0)$ we simply apply the functions--sheaves dictionary
  as follows ({\it cf} \cite{laumon:87a}*{1.1.3} or, ultimately, \cite{deligne:SGA4.5}*{Sommes trigonométriques}).
  First, recall that $L_G(f,\psi) = (L(f,\psi), \mu(f,\psi))$.
  Let $k'/k$ be any degree-$n$ extension and let $\Nm : k'\to k$ be the norm map. 
  By \cite{laumon:87a}*{1.1.3.3}, 
  \[
  \trFrob{(G_{k'}\hookrightarrow G)^*L(f,\psi)} = \trFrob{\qcs{L}} \circ \Nm;
  \]
  on the other hand, 
  \[
  \trFrob{\qcs{L}} \circ \Nm = \trFrob{(G_{k'}\hookrightarrow G)^*\qcs{L}},
  \]
   by Proposition~\ref{prop:basechange}.
  %Using notation from \cite{laumon:87a}*{1.1}, this means $t_{\bullet L(f,\psi)} = t_{\bullet \qcs{L}_0}$.
  Using \cite{laumon:87a}*{Th\'eor\`eme~1.1.2}, it follows that  $L(f,\psi) \iso \qcs{L}_0$ in $\Loc(G)$.
  It is now clear that $(L(f,\psi),\mu(f,\psi)) \iso (\qcs{L}_0,\mu_0)$ in $\QCb(G)$.
  This completes the proof that $L_G : \QCf(G) \to \QC(G)$ is essentially surjective.
  Since we have already seen (in Section~\ref{sec:finite}) that $L_G$ is full and faithful, 
  it follows that $L_G$ is an equivalence.

@@

Recall that $\Fq$ is a finite field.
Let $K$ be a non-archimedean local field with residue field $\Fq$.
We place no restriction on the characteristic of $K$. 
We write $\OK$ for the ring of integers of $K$ and $\pK$ for the maximal ideal in $\OK$. 
%For any positive integer $d$, set $A_d\ceq \OK/\pK^d$.
Let $T$ be an algebraic torus over $K$.
We write $\Hom_{< d}(T(K),\EEx)$ for the group of characters of $T(K)$ depth less than $d$ 
and $\Hom_{<\infty}(T(K),\EEx)$ for the group of characters of $T(K)$ of arbitrary finite depth; one might call these quasicharacters of $T(K)$, expanding on \cite{tate:67a}*{2.3}.

@@

\begin{definition}\label{def:GN}
For each positive integer $d$, set $\TT_{d} \ceq \TT \times_\Spec{\OK} \Spec{\OK/\pK^d}$. Observe that $\OK/\pK^d$ is an Artin ring and $\TT_{d}$ is smooth and locally of finite type over $\OK/\pK^d$.
%
Let $\GN{T}_{d} = \Gr(\TT_d)$ be the group scheme over $\Fq$ produced by applying
the Greenberg functor to $\TT_{d}$; this is the \emph{Greenberg transform of $\TT_d$}.
\end{definition}

See \cite{greenberg:63a} and \cite{bosch-lutkebohmert-reynaud:NeronModels}*{Ch. 9, \S 6} and the first few sections of \cite{stasinski:12a} for background on the Greenberg functor.

For each positive integer $d$, consider the canonical surjective ring homomorphisms $\OK/\pK^{d+1} \to \OK/\pK^d$ and the associated cartesian diagram of schemes 
\[
\begin{tikzcd}
\Spec{\OK/\pK^d} \arrow[hook]{rr} && \Spec{\OK/\pK^{d+1}} \arrow[hook]{r} & \Spec{\OK} \\
 \arrow{u} \TT_{d} \arrow[hook]{rr} & \ \arrow[Rightarrow]{d}{\text{Greenberg functor}} & \arrow{u} \TT_{d+1}  &  \\
\GN{T}_{d} \arrow[hook]{rr} & \ & \GN{T}_{d+1}  \arrow{r} & \Spec{k}
\end{tikzcd}
\]
The map $\GN{T}_{d} \hookrightarrow \GN{T}_{d+1}$ is not a morphism of group schemes. There is, however, an epimorphism of group schemes $\GN{T}_{d+1} \to \GN{T}_{d}$ for which $\GN{T}_{d} \hookrightarrow \GN{T}_{d+1}$ is a section.

The fundamental properties of $\GN{T}_d$ that we need are:
 \begin{enumerate}
   \labitem{(GN.0)}{dFq} $\GN{T}_d(\Fq) = \TT(\OK/\pK^d)$, canonically.
  \labitem{(GN.1)}{GNd} $\GN{T}_d$ is a smooth, commutative group scheme, locally of finite type over $\Fq$, and the component group of $\GN{T}_d$ is geometrically finitely generated. (In fact, $\pi_0(\GN{T}_d) = \pi_0(\TT) \times_{\Spec{\OK}} \Spec{\Fq}$, and the latter is finitely generated by \cite{xarles:93a}.)
  \labitem{(GN.2)}{d+1} 
  \end{enumerate}

@@

\begin{proof}
  By \ref{GNd}, we may apply Theorem~\ref{thm:snake} to
  $\GN{T}_d$ which provides a canonical isomorphism
  $\QCiso{\GN{T}_d} \iso \GN{T}_d(\Fq)^*$.
  By \ref{dFq}, $\GN{T}_d(\Fq) = \TT(\OK/\pK^d)$.
  Then we recall that the filtration of $T(K)$ used to
  define depth has the property that $\TT(\OK/\pK^d) \cong T(K) / T(K)_d$
  \cite{yu:03a}*{Prop 5.2}, so characters of $\GN{T}_d(\Fq)$ 
  are in bijection with characters of $T(K)$ vanishing on $T(K)_d$. 
  This shows that $\trFrob{T,d}$ is an isomorphism.

@@
